---
title: "EXPLORATORY DATA ANALYSIS : RETAIL"
author: "Objectives: To Perform Exploratory Data Analysis on dataset SampleSuperstore. As a business manager, try to find out the weak areas where you can work to make more profit. What all business problems you can derive by exploring the data."
date: "Author: Irshad"
output: html_document
---

## Installing and loading necessary libraries
```{r message=FALSE, warning=FALSE}
install.packages("tidyverse")
library(tidyverse)
install.packages("ggplot2")
library(ggplot2)
install.packages("skimr")
library(skimr)
install.packages("janitor")
library(janitor)
install.packages("dplyr")
library(dplyr)

```

## Importing the dataset

```{r message=FALSE, warning=FALSE}
Superstore <- read_csv("SampleSuperstore.csv") #load the data
```

## Dataset features

```{r pressure, echo=TRUE, message=TRUE}
head(Superstore) ## Prints the top 6 rows of data
```
```{r include=FALSE}
Superstore <- rename(Superstore,Shipmode = "Ship Mode")
```
```{r include=FALSE}
Superstore <- rename(Superstore,Subcategory = "Sub-Category")
```


## Information about the dataset

```{r}
skim_without_charts(Superstore)
```

```{r}
glimpse(Superstore)
```


```{r}
colnames(Superstore) ##Displays the column names
```


```{r}
sum(is_null(Superstore)) ## Checking for null values 
```

#### Checking if there are any duplicate rows


```{r}
sum(duplicated(Superstore)) 
```

```{r}
Superstore <- unique(Superstore) ##Removing Duplicate
```



```{r}
str(Superstore) ##Duplicates removed 
```

## Checking for Outliers 

```{r}
ggplot(Superstore) +
  aes(x = Discount) +
  geom_histogram(bins = 30L, fill = "#0a4b8a") +
  theme_minimal()
```

```{r}
Superstore <- filter(Superstore, Discount > 0.45) ##filtering outliers 
```


```{r}
head(Superstore$Discount) #filtered outliers
```
## Data visualization

### * Analysing Shipmode

```{r}
ggplot(Superstore, aes(Shipmode)) + geom_bar(aes(fill = Shipmode))
```

### Analysis 1
#### we can find out that standard class ship mode is preferred more than the other ship modes are available.

## * Analysing Segment

```{r}
ggplot(Superstore, aes(Segment )) + geom_bar(aes(fill = Segment )) +  scale_fill_brewer(palette = "Blues") 
```

### Analysis 2
#### Consumer has the highest count than corporate and home office.

## * Analysing Category

```{r}
ggplot(Superstore, aes(Category)) + geom_bar(aes(fill = Category)) + scale_fill_grey( 
start = 0.2, end = 0.8, 
na.value = "red") 
```

### Analysis 3
#### Office Supplies Category has the highest count.

## * Analysing Region

```{r}
ggplot(Superstore, aes(Region)) + geom_bar(aes(fill = Region)) + scale_fill_brewer(palette = "Reds")
```

### Analysis 4
#### Here West region has the highest customer's count

```{r include=FALSE}
Superstore <- read_csv("SampleSuperstore.csv")
```
```{r include=FALSE}
Superstore <- rename(Superstore,Shipmode = "Ship Mode")
Superstore <- rename(Superstore,Subcategory = "Sub-Category")
```

## * Analysing Category and Sub-Category

```{r}
ggplot(data=Superstore) + geom_bar(mapping = aes(x=Category , fill=Subcategory)) + facet_wrap(~ Subcategory) +  theme(axis.text.x = element_text(angle = 30))
```

### Analysis 5
#### * Furniture which consists of 4 subcategories items Bookcases, Chairs, Tables, Furnishings
#### * Office Supplies category consist 9 sub-categories Labels, Storage, Art, Binders, Appliances, Paper, Envelopes, Fasteners, Supplies
#### * Technology involves 4 subcategories, those are Phones, Accessories, Machines, Copiers.

```{r include=FALSE}
Superstore <- filter(Superstore, Discount > 0.45)
```

## * Analysing Profit, Discount and Sales
```{r}
ggplot(Superstore, aes(Profit , Sales)) + 
 geom_point(aes(color = Discount , size=Discount)) + 
 geom_smooth(method ="lm") + 
 coord_cartesian() + 
 scale_color_gradient() + 
 theme_bw()
```

## * sales vs profit

```{r}
ggplot(data=Superstore, aes(x=Sales,y=Profit,col= Shipmode, shape=Shipmode))+ geom_point()+geom_smooth(method ="lm")+labs(title="Sales vs Profit")
```

### Analysis 6
#### More profit/Loss from Standard class.

## * Sales vs Discount

```{r}
ggplot() + geom_point(data=Superstore, aes(x=Discount,y=Sales,col= Shipmode, shape=Shipmode))+labs(title="Sales vs Discount")
```

### Analysis 7
#### Discount attracts mostly the Standard class ,Same day receive least Discount

## * Category vs Discount

```{r}
ggplot() + geom_bar(data=Superstore, aes(x=Category,y=Discount,fill= Region),stat = "identity")
```

### Analysis 8
### office suppliers gets more Discount 


## * Profit and Discount trade off in the different category

```{r}
ggplot(Superstore, aes(Profit ,Discount )) +   geom_jitter(aes(shape = Category, color = Category)) + facet_grid(~ Category) 
```

### Analysis 9
#### As discount is increased 0.5 or more losses are faced in all categories product.


## *  Sales and discount of Product effect in the different category product.

```{r}
ggplot(Superstore, aes(Sales ,Discount )) +   geom_jitter(aes(shape = Category, color = Category))  +facet_wrap(~ Category)+ theme(axis.text.x = element_text(angle = 20))

```

### Analysis 10
#### * maximum discount on 0.7 and 0.8, Sales of Office supplies was increased.
#### * Sales of Furniture category was increased due to discount.
#### * Sales of Technology category remains same. 



## Conclusion

#### * Discounts should be based on sales and should not increase after a particular range otherwise unnecessary   discounts with low sales can witness huge losses.
#### * Binders and Machine industry should be focused upon more, Binders gets more discount but gives more losses   Office Suppliers sales less but gets more discount and gives losses
#### * Central Region should not offer any kind of discounts for better performance.
#### * Furniture Category performance is less comparing to others so discount should not be give to them.
#### * If Same day shipment receives more discounts then can trigger more sales/Profit.



## THANK YOU





